
```{r, echo=FALSE, include=FALSE}
### SET OPTIONS
#It looks like overwrite=TRUE in the file.copy line was necessary
library(knitr)
opts_chunk$set(fig.path='figure/minimal-', fig.align='center', 
               fig.show='hold', concordance=TRUE, external=TRUE,
               tidy=TRUE, size='footnotesize', verbose=TRUE,
               purl=TRUE, warning=FALSE, echo=FALSE)
```

# HIVBackCalc: Analysis Report
 _`r format(Sys.time(), "%d %B, %Y")`_


## Raw Data

### eHARS?

```{r}
    if (is.null(input$ehars)) {
        eh <- 'are NOT'
    } else if (!input$ehars) {
        eh <- 'are NOT'
    } else eh <- 'ARE'
```
The input data `r eh` raw eHARS data.

### Data preview
The first 10 rows and 20 or so columns are displayed below.

```{r}
    ncol <- ncol(rawdata())
    d1 <- ifelse(ncol>10, 10, ncol)
    ################################################## 
    # LOAD AND DISPLAY DATA
    ################################################## 
    kable(rawdata()[1:10,1:d1])
```
```{r}
    if (ncol>10) kable(rawdata()[1:10,(d1+1):ncol])
```

### Data diagnostics 
```{r}
    ################################################## 
    # CHECK DATA
    ################################################## 
    checkData()
```

### Summary of raw data

*Numeric variables*
```{r}
    kable(formattedDataList()$rawVarSum$Numeric,
        caption='Summary of numeric variables in the data, pre-formatting',
        label='tab:format2',
        digits=0,
        caption.placement='top',
        include.rownames=TRUE,
    )
```

*Categorical variables*
```{r}
    kable(formattedDataList()$rawVarSum$Categorical,
        caption='Summary of non-numeric variables in the data, pre-formatting. The display of values is truncated after 25 unique values.',
        label='tab:format2',
        digits=0,
        caption.placement='top',
        include.rownames=FALSE,
        size='small'
        )
```

### Summary of data formatting
```{r}
    ################################################## 
    # FORMAT DATA
    ################################################## 
    kable(formattedDataList()$assumptions[,c(1,3)],
        caption='Formatting assumptions applied to the data, in addition to the assumption selected for those with no testing history. Missing month/day and illogical last negative are relevant only when the input data are raw eHARS data.',
        label='tab:format1',
        digits=0,
        caption.placement='top',
        include.rownames=FALSE,
        size='small'
        )
```

## Analysis selections/features
The remainder of the report refers to the data given the selections indicated below.

### Years

```{r}
if (is.null(input$selectedYears)) {
    years <- c(min(floor(dataf()$timeDx)), max(floor(dataf()$timeDx)))
} else years <- input$selectedYears
```

The years selected are `r years[1]` to `r years[2]`.

### Subgroups and stratification
```{r}
    ################################################## 
    # SELECT SUBSET
    ################################################## 
```

+ Analysis (sub)group: `r subgroupVar()`
+ Stratified by: `r stratVar()`

### Supplemental data on diagnosed People Living with HIV (PLWH)

```{r}
    ################################################## 
    # PLWH
    ################################################## 
    kable(plwh())
```

## Data Exploration

### Sample characteristics and testing history
```{r}
    ################################################## 
    # DESCRIBE SAMPLE
    ################################################## 
    dataf <- dataf()
    variables <- c(`Age Group`='agecat5', 
                   `Race/Ethnicity`='race', 
                   `Mode of Transmission`='mode')

    everHadNegTest_subgrouptab <- tabTestHist(dataf, variables, 
                                              supercolumn=TRUE,
                                              fullsample_row=TRUE)
    kable(everHadNegTest_subgrouptab,
        caption='Column % sums to 100 within each characteristic. Availability of testing history data within each subgroup level is shown as row percents of %Yes, %No, and %Missing',
        label='tab:sample',
        digits=0,
        table.placement='!h',
        caption.placement='top',
        include.rownames=FALSE,
        size='small',
        sanitize.text.function=function(str) { gsub('(\\.)*Percent(\\.)*', ' \\% ',str); }
        )
```

### Testing histories over time
```{r, fig.height=3}
    ################################################## 
    # PLOT TESTING HISTORIES
    ################################################## 
    plotTestHist(dataf())
```

### Diagnosis counts over time
```{r, fig.height=3}
plotDiagnoses(dataf())
```

```{r}
      dataf <- dataf()
      if (is.null(stratVar())) {
      t1 <-     table(dataf$timeDx)
      } else if (stratVar()=='None') {
      t1 <-     table(dataf$timeDx)
      } else {
      t1 <-     table(dataf[,stratVar()], dataf$timeDx)
      }
      kable(t(t1),
     caption='Diagnoses per time step, by strata',
     label='tab:dxByMSM',
     digits=0,
    table.placement='!h',
    caption.placement='top',
    include.rownames=TRUE,
    size='small',
    sanitize.text.function=function(str) { gsub('(\\.)*Percent(\\.)*', ' \\% ',str); })
```

## Time from infection to diagnosis (TID)
```{r, fig.height=4}
plot(TIDs(), intLength=diagInterval(), 
     cases = c('Base Case', 'Upper Bound'))
```
```{r}
    pid <- TIDs()$base_case$cdf
    pidU <- TIDs()$upper_bound$cdf

    # Survivor fxn by quarter-year
    pdf_dataframe <- data.frame(yrs=c(0, seq(diagInterval(),
                                             by=diagInterval(),
                                             length.out=length(pid))),
                                surv=c(1, 1-pid),
                                survU=c(1, 1-pidU))
    # Focus on years and half-years, otherwise there's too much info
    tidtimes <- pdf_dataframe$yrs
    pdf_dataframe <- pdf_dataframe[round(tidtimes*2)==tidtimes*2,]

    colnames(pdf_dataframe) <- c('Years since infection',
                                 'Base Case',
                                 'Upper Bound')
    rownames(pdf_dataframe) <- NULL

    kable(pdf_dataframe,
          include.rownames=FALSE, 
          caption='Fraction still undiagnosed under the two cases',
          caption.placement='top',
          digits=3)
```

## Time step and sample size

The time step defining the time of diagnosis is `r as.numeric(diagInterval())` year(s). This results in the following distribution of diagnoses per time step:

```{r}
          dataf <- dataf()
          diagBreaks <- c(0,5,20,50,999)
          if (is.null(stratVar())) {
              diagCounts = tabulateDiagnoses(dataf, intLength=diagInterval())
              t2 <- table(cut(diagCounts[!is.na(diagCounts)], 
                        breaks=diagBreaks,
                        include.lowest=TRUE,
                        right=FALSE))
          } else if (stratVar()=='None') {
              diagCounts = tabulateDiagnoses(dataf, intLength=diagInterval())
              t2 <- table(cut(diagCounts[!is.na(diagCounts)], 
                        breaks=diagBreaks,
                        include.lowest=TRUE,
                        right=FALSE))
          } else {
              t2 <- ddply(dataf, c(stratVar()), .fun=function(x) {
                    diagCounts = tabulateDiagnoses(x, intLength=diagInterval())
                    table(cut(diagCounts[!is.na(diagCounts)], 
                              breaks=diagBreaks,
                              include.lowest=TRUE, right=FALSE)) 
                             })
          }

          kable(t2,
                size='small',
                caption='Number of time steps having 0-4, 5-19, 20-49, and 50-999 diagnoses, by selected stratum variable',
                caption.placement='top')
```

## Results

### Diagnoses, incidence and undiagnosed counts

*Summary over all time steps*
```{r}
kable(results()$resultsSummary,
      caption='Summary of reported diagnoses, estimated incidence for the two TID cases, and estimated undiagnosed counts for the two TID cases across all time periods.', 
      caption.placement='top'
      )
```

*By time step*
```{r, fig.height=5}
plot(results())
```

*By year - see next table*

### True prevalence and undiagnosed fractions (if PLWH available)

```{r, fig.height=5}
if (!is.null(trueprev())) trueprev()$plot2
```

```{r, fig.height=5}
tFull <- tableToDownload()[,c('Estimate', 'Diagnoses/Case', 'Year', 'Median', 'Mean')]
tFull <- tFull[order(tFull$Estimate),]
rownames(tFull) <- NULL
kable(tFull, 
      include.rownames=FALSE,
      caption='Results by year', 
      caption.placement='top')
```


